Scalar Items in Embedded Position: {A}n Experimental Revisit

### 2012-07-20 text/paper/paper.tex

Summing up so far, three previous studies have presented different
pieces of evidence concerning putative local
readings. \citeauthor{GeurtsPouscoulous2009:Embedded-Implic}'s results
suggest that local readings are not available, because they do not
affect a categorical truth-value judgments in a picture-verification
task. On the other hand,
\citeauthor{CliftonDube2010:Embedded-Implic}'s and
\citeauthor{ChemlaSpector2010:Experimental-Ev}'s design was suggested
to be sensitive enough to reveal effects of local readings: either by
a choice of the situation that better befits a given sentence
(\citeauthor{CliftonDube2010:Embedded-Implic}), or by a graded
truth-value judgment
(\citeauthor{ChemlaSpector2010:Experimental-Ev}).

The data made available by all of these three studies, if taken in
conjunction, can be interpreted in several ways. Here are the two most
obvious candidates that have been suggested in the literature.
Firstly, conventionalists or grammaticalists could maintain that the
latter two studies refute traditionalism because they show that local
readings are available, despite the fact that they do not affect
simple \mymark{categorical truth-value judgements}. This is
essentially the interpretation favored by
\citet{CliftonDube2010:Embedded-Implic}, as well as
\citet{ChemlaSpector2010:Experimental-Ev}. In contrast,
traditionalists could rebut this position by arguing that the very
reason why local readings do not affect categorical judgements is
because they are not truth-relevant speaker-intended meaning
enrichments after all. The traditionalist, who would like to argue in
this way, would then have to provide an alternative explanation for
the effects attributed to the availability of local readings in the
latter two studies. One such explanation, at least for the data
concerning \as-sentences, has recently been suggested by
\citet{Tielvan-Tiel2012:Embedded-Scalar}, who argues that the effects
of putative local readings in the latter mentioned studies are due to
\mymark{typicality effects}.\fn{Yet another possible interpretation
  could be given by literalism which admits to local readings but
  considers them dispreffered to literal readings. This position
  explains the tendentious effects reported in the latter two studies
  \emph{and} explains why local readings do not show in categorical
  judgements: literal readings are simply preferred, so that local
  readings only figure in tendentious choices, but not in categorical
  ones. We will come back to this position eventually.}

@@

However, the interpretation of data from judgement tasks is generally
problematic for reasons already addressed by
\citet{ChemlaSpector2010:Experimental-Ev} and, more recently,
\citet{Tielvan-Tiel2012:Embedded-Scalar}. The problem is that the
present variants of truth-value judgment tasks all run risk of
conflating evidence about the availability or preference of readings
with properties of the pictorial material that are independent of the
reading to be assessed. This problem affects all of the studies
covered above, to some degree or other, because it follows almost
directly from the nature of the phenomenon under investigation and
therefore seems nearly impossible to avoid.

>>>>

Summing up so far, three previous studies have presented diverging
pieces of evidence concerning the status of putative local
readings. \citet{GeurtsPouscoulous2009:Embedded-Implic} did not observe any truth value judgments that would indicate
local readings. On the basis of this finding they argue that these readings are simply not available. On the other
hand, C\&D and C\&S argue that the findings of G\&P may be due to a bias in their experimental design and
employed different methods intended to be sensitive enough to reveal effects of local readings: either by
a choice of the situation that better befits a given sentence (\citeauthor{CliftonDube2010:Embedded-Implic}), or by a graded
truth-value judgment (\citeauthor{ChemlaSpector2010:Experimental-Ev}). Both studies found evidence in support of local readings.

To account for all of these observations, it could be hypothesized
along with \citeauthor{CliftonDube2010:Embedded-Implic} and
\citeauthor{ChemlaSpector2010:Experimental-Ev} that local readings are
available, but so dispreferred that they do not affect
\emph{categorical} truth-value judgements (See, however, \citet{Crain1998} for arguments in favour of truth value judgments
as a means of detecting highly dispreferred readings).
It could then furthermore be argued, in line with traditionalism, that the reason why local
readings do not affect categorical judgements is because they are not
(strongly) truth-relevant speaker-intended meaning enrichments after
all, so that no case can be made for lexicalism or grammaticalism
\citep[c.f.][for arguments along these lines]{Tielvan-Tiel2012:Embedded-Scalar}.

@@

These observed patterns suggest that local readings \emph{are}
available and \emph{can} have an effect even on categorical
truth-value judgements.
